*Salmonella* Environmental Persistence Informs Management Relevant to Avian and Public Health

### 2023-04-26 3. Analysis Script/StatisticalAnalysis.qmd
Auto stash before merge of "master" and "origin/master"

lmfit_PT<- lmog %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Date, PT_Data)

@@

tidy(lmfit_PT) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))

#Fitting other specific indicators
lmog_1<-linear_reg() %>% 
  set_engine("lm")

lmfit_PT1<- lmog_1 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Sample_Type+Date, PT_Data)

@@

@@

tidy(lmfit_PT1) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))

#Trying ALL predictors of Picnic Table dataset 
lmog_PT<- linear_reg() %>%
  set_engine("lm")

@@

lmfit_PTglm2<- lmog_glm2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Quadrant+Table, PT_Data)

lmfit_PTglm2

@@

tidy(lmall_PTGLM)
```
# **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
```{r}
PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial)

@@

PT_GLM_PREV
```
# **Generalized Linear Model: Feeder Data**
```{r}
lmog_FD<-linear_reg() %>% 
  set_engine("glm")

@@

FD_GLM_PREV
```
#**Time to event**
```{r}
library(finalfit)
library(forcats)
library(survival)

@@

as.Date(PT_Data$Date, format= "%Y-%M-%D")

@@

PT_Data = PT_Data %>% 
  mutate(
    Salmonella_Positive_os = ifelse(Salmonella_Positive == 0,1,0),
      status_crr= case_when(
        `Salmonella_Positive` == 1~0,
        `Salmonella_Positive` == 1~1))

@@

PT_Data %>% 
   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
    rename("Overall survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all) 
```

>>>>

lmfit_PT<- model_lm %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Date, PT_Data)

@@

tidy(lmfit_PT) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
```

```{r}
#Fitting other specific indicators
lmog_1<-linear_reg() %>% 
  set_engine("lm")

model_lm1<- linear_reg()

lmfit_PT1<- model_lm1 %>% #Training; outcome: Salmonella Pos
  fit(Salmonella_Positive~Sample_Type, PT_Data)

@@

tidy(lmfit_PT1)

@@

tidy(lmfit_PT1) %>%
    dwplot(dot_args = list(size = 2, color = "coral1"),
         whisker_args = list(color = "coral1"),
         vline = geom_vline(xintercept = 0, color = "darkred", linetype = 4))
```

```{r}
#Trying ALL predictors of Picnic Table dataset 
lmog_PT<- linear_reg() %>%
  set_engine("lm")

@@

lmfit_PTglm2<- lmog_glm2 %>% #Training; outcome: Salmonella Pos and Main Predictor Table
  fit(Salmonella_Positive~Quadrant+Table, PT_Data) 
```

```{r}
lmfit_PTglm2

@@

tidy(lmall_PTGLM)
```

# **Generalized Linear Model: Picnic Table (Persistence & Prevalence)** 
```{r}
PT_GLM_PER <- glm(Salmonella_Positive ~ Date, data = PT_Data, family = binomial)

@@

PT_GLM_PREV
```

# **Generalized Linear Model: Feeder Data**
```{r}
lmog_FD<-linear_reg() %>% 
  set_engine("glm")

@@

FD_GLM_PREV
```
# **Time to event**
```{r}
library(finalfit)
library(forcats)
library(survival)

@@

@@

PT_Data = PT_Data %>% 
  mutate(
      salmonella_crr= case_when(
        Salmonella_Positive ==  0~1
        Salmonella_Positive == 1~1
        TRUE ~ 1))

@@

PT_Data %>% 
   finalfit(dependent_crr, explanatory, add_dependent_label = FALSE) %>% 
    rename("Overall survival" = label) %>% 
    rename(" " = levels) %>% 
    rename("  " = all)

survfit(Surv(Date, Salmonella_Positive)~1, data=PT_Data)

PT_Data %>%
  filter(status==1) %>%
  summarize(median_surv=median(Date))
```

### 2023-04-26 5. Manuscript/Manuscript.qmd
Analysis Edits continued...
Auto stash before merge of "master" and "origin/master"

![Positive Salmonella Samples by Type and
Date](images/Picnic_Table_PosSamp-04.png) Positive *Salmonella* Samples
by Sample Type and Date

![Positive *Salmonella* Samples by Quadrant and
Date](images/Picnic_Table_Quad-01.png)

![Positive *Salmonella* Samples by feeder type and
date.](images/Picnic_Table_PosSamp.png)

@@

@@

### **Persistence**
On Day 1 of the study, four fecal piles per table (one per quadrant, n = 12) were designated for tracking *Salmonella* persistence.  **Salmonella** was isolated from three of the 12 designated persistence study piles on Day 1, allowing the persistence to be tracked on these three piles.   Persistence, expressed as the number of days that a positive fecal pile on Day 1 remained positive, ranged from 1 to 4 days with an average persistence of 3 days. The probability of an already positive sample to remain positive the next day declined consistently throughout the study period.

### **Prevalence**
Of 96 deposited feces, we obtained 27% positive. *Salmonella* persisted for an average of 3 days, and maximum of 4 days on wooden picnic tables. *Salmonella* was isolated from all tables, at least once. While there appears to be a significant positive correlation between the number of new fecal piles and the probability of yielding a positive pooled sample. The probabilities of a positive sample varied throughout the week, with day 4 having the lowest probability of positive *Salmonella* isolation and day 7 having the most error.

## **Feeder Trial**
We confirmed successful inoculation of all feeders with *Salmonella*
(presence on day post inoculation \[DPI\] 0). Further, results indicate
that *Salmonella* may persist on some feeders up to DPI 5 and current
analysis indicates that *Salmonella* persisted on both coated and
uncoated plastic feeders at a higher rate than wood and wood coated.
Figure 2 depicts the percent of samples positive based on feeder type
throughout the study period. While plastic and plastic coated feeders
had an overall higher percent positivity, a spike in percent positivity
in wood feeders occurred on DPI 5 after a rain event. Further analyses
should be conducted to determine whether environmental conditions (e.g.,
rain event) had a significant influence on persistence.

>>>>

@@

![Positive Salmonella Samples by Type and
Date](images/Picnic_Table_PosSamp-05.png) Positive *Salmonella* Samples
by Sample Type and Date

![Positive *Salmonella* Samples by Quadrant and
Date](images/Picnic_Table_Quad-02.png)

![Positive *Salmonella* Samples by feeder type and
date.](images/Feeder_PA.png)

@@

### **Persistence**

On Day 1 of the study, four fecal piles per table (one per quadrant, n =
12) were designated for tracking *Salmonella* persistence.
**Salmonella** was isolated from three of the 12 designated persistence
study piles on Day 1, allowing the persistence to be tracked on these
three piles. Persistence, expressed as the number of days that a
positive fecal pile on Day 1 remained positive, ranged from 1 to 4 days
with an average persistence of 3 days. The probability of an already
positive sample to remain positive the next day declined consistently
throughout the study period.

### **Prevalence**

Of 96 deposited feces, we obtained 27% positive. *Salmonella* persisted
for an average of 3 days, and maximum of 4 days on wooden picnic tables.
*Salmonella* was isolated from all tables, at least once. While there
appears to be a significant positive correlation between the number of
new fecal piles and the probability of yielding a positive pooled
sample. The probabilities of a positive sample varied throughout the
week, with day 4 having the lowest probability of positive *Salmonella*
isolation and day 7 having the most error.

## **Feeder Trial**

We confirmed successful inoculation of all feeders with *Salmonella*
(presence on day post inoculation \[DPI\] 0). Further, results indicate
that *Salmonella* may persist on some feeders up to DPI 5 and current
analysis indicates that *Salmonella* persisted on both coated and
uncoated plastic feeders at a higher rate than wood and wood coated.
Figure 2 depicts the percent of samples positive based on feeder type
throughout the study period. While plastic and plastic coated feeders
had an overall higher percent positivity, a spike in percent positivity
in wood feeders occurred on DPI 5 after a rain event. Further analyses
should be conducted to determine whether environmental conditions (e.g.,
rain event) had a significant influence on persistence.
